An Algorithm for Searching Evolving Graphs in Linear Algebraic Language

### 2016-01-21 paper.tex
Rewrite description of algebraic formulation

There exits a duality between vector matrix multiplication and the BFS
for graphs \cite{kegi11}. Let $G = (V, E)$ be a (directed or undirected) graph
and $k$ a node in the graph $G$. We could represent $G$ as an adjacency
matrix $A$ and $k$ a column vector $b$ such that
\[
b_i =
\begin{cases}
1 & \mbox{if $i = k$,} \\
0 & \mbox{otherwise.}
\end{cases}
\]
Then $A^Tb$ is a column vector where the indices of the nonzero entries
indicate the forward-neighbors of $k$. This relationship can be generalized to be
a duality between a block vector matrix multiplication and the
evolving-graph BFS.

For the evolving case, we shall introduce a new binary operator $\odot$.
We define
\[
A \odot b =
\begin{cases}
b & \mbox{if $Ab \ne 0$,} \\
0 & \mbox{otherwise,}
\end{cases}
\]
where $A$ is an adjacency matrix and $b$ a column vector defined as before.
Then it is easy to observe that
the forward-neighbors of node $(k, 1)$ in $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ are the nonzero indices of the
elements in the sequence
\begin{equation}
\label{eq:out_ne}
\big\langle (A^{[1]})^Tb, \quad  (A^{[2]})^T\odot b, \quad \ldots \quad (A^{[n]})^T\odot b \big\rangle.
\end{equation}
The forward-neighbors at timestamp $t$ are the nonzero indices of the $t$-th element of the above sequence.
Hence given $A_n$ and a vector $b$, we can use \eqref{eq:out_ne} to
get a $1$-step  BFS tree. In Figure \ref{fig:eg_shortest_path}, the forward-neighbors
of node $(1, t_1)$ can be computed by
\[
\Big\langle
  \begin{bmatrix}
    0 & 0 & 0 \\
    1 & 0 & 0 \\
    0 & 0 & 0
  \end{bmatrix}
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
,
 \begin{bmatrix}
   0 & 0 & 0 \\
   0 & 0 & 0 \\
   1 & 0 & 0
 \end{bmatrix}
\odot
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
,
 \begin{bmatrix}
  0 & 0 & 0 \\
  0 & 0 & 0 \\
  0 & 1 & 0
 \end{bmatrix}
\odot
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
\Big\rangle
\]
which is equal to
\[
\Big\langle
\begin{bmatrix}
0 \\
1 \\
0
\end{bmatrix},~
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix},~
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix}
\Big\rangle
\]
This means node $(2,t_1)$ and $(1,t_2)$ are the forward-neighbors of $(1,t_1)$.

>>>>

This algebraic formulation exploits a graphical interpretation of matrix--vector
products involving the adjacency matrix~\cite{kegi11}.
Let $G = (V, E)$ be a (directed or undirected) graph
and $k$ be a node in the graph $G$. If $A$ is the adjacency matrix of $G$ and
$b$ is the column vector
\[
b_i =
\begin{cases}
1 & \mbox{if $i = k$,} \\
0 & \mbox{otherwise,}
\end{cases}
\]
then $A^Tb$ is a column vector where the indices of the nonzero entries
indicate the forward-neighbors of $k$.

The evolving graph BFS can be thought of as a generalization of blocked
matrix--vector products.
Introduce a new operator $\odot$, a variation on the ordinary matrix--vector
product such that
\[
A \odot b =
\begin{cases}
b & \mbox{if $Ab \ne 0$,} \\
0 & \mbox{otherwise.}
\end{cases}
\]
Then, the forward-neighbors of node $(k, 1)$ in $A_n = \langle A^{[1]}, A^{[2]}, \ldots, A^{n]}\rangle$ are the nonzero indices of the
elements in the sequence
\begin{equation}
\label{eq:out_ne}
\big\langle (A^{[1]})^Tb, \quad  (A^{[2]})^T\odot b, \quad \ldots \quad (A^{[n]})^T\odot b \big\rangle.
\end{equation}
The forward-neighbors at timestamp $t$ are the nonzero indices of the $t$-th element of the above sequence.
Hence given $A_n$ and a vector $b$, we can use \eqref{eq:out_ne} to
get a $1$-step BFS tree. In Figure \ref{fig:eg_shortest_path}, the forward-neighbors
of node $(1, t_1)$ can be computed by
\[
\Big\langle
  \begin{bmatrix}
    0 & 0 & 0 \\
    1 & 0 & 0 \\
    0 & 0 & 0
  \end{bmatrix}
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
,
 \begin{bmatrix}
   0 & 0 & 0 \\
   0 & 0 & 0 \\
   1 & 0 & 0
 \end{bmatrix}
\odot
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
,
 \begin{bmatrix}
  0 & 0 & 0 \\
  0 & 0 & 0 \\
  0 & 1 & 0
 \end{bmatrix}
\odot
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix}
\Big\rangle
\]
which is equal to
\[
\Big\langle
\begin{bmatrix}
0 \\
1 \\
0
\end{bmatrix},~
\begin{bmatrix}
1 \\
0 \\
0
\end{bmatrix},~
\begin{bmatrix}
0 \\
0 \\
0
\end{bmatrix}
\Big\rangle
\]
This means node $(2,t_1)$ and $(1,t_2)$ are the forward-neighbors of $(1,t_1)$.
